---
title: "Homework 3 Q1"
author: "Wonhee Lee, Eugene Kang"
date: "September 13, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load crime data

```{r data}
crime <- read.table("uscrime.txt", header = TRUE)
head(crime)
```


## Plot crime data to get a sense of the distribution of data points

```{r crime}
plot(crime$Crime)
```

## Install & load outliers package

```{r package}
library(outliers)
```

## Perform grubbs.test to check for outliers. For the purpose of this analysis, we will assume an alpha of .05.
### Check for highest value(1993)
```{r high}
high = grubbs.test(crime$Crime, type=10)
high
high$p.value < .05
```
### Since p-value is greater than alpha, we fail to reject the null hypothesis that 1993 is not an outlier.

### Check for lowerst value(342)
```{r low}
low = grubbs.test(crime$Crime, type=10, opposite = TRUE)
low
low$p.value < .05
```
### Since p-value is greater than alpha, we fail to reject the null hypothesis that 342 is not an outlier.